---
title: "Week2"
author: "Thomas Rosenthal"
date: "21/01/2021"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir  = '../..')
```

### Shiny Functionality
I've spent most of this week playing around with some of the basic functionality of Shiny. 

In _Mastering Shiny_, Wickham lays out the "Getting Started" section into three categories:
- The basics: to help get us going with Shiny
    - the simplest possible ```hello world``` app
    - a few simple plots, some sliders and other input formats 
    - and an overview of Shiny's general capabilities.
- Some structural UI necessities: to help get a feel for laying out the app.
- Reactivity: to help to lay the foundation for _how_ shiny works, at least from a purely functional perspective

I did this chapter after building the [mpg app](https://shiny.rstudio.com/articles/build.html), where I essentially wrote out the code (rather than copying and pasting, which helped immensely) and changed some colours around. I was glad I did it in this order actually. My version of this app (though it's 99% the same) is [here](https://github.com/mrpotatocode/COFEE_COFFEE_COFFEE/tree/main/Tamatoa/First). 

### Exercises

For [Chapter 2](https://mastering-shiny.org/basic-app.html#exercises) I did a total of five exercises.

__2.8.1:__
```{r}
library(shiny)
ui <- fluidPage( 
    
    # Application title
    titlePanel("Chapter 2 Exercises"),
    
        # Main panel for displaying outputs 
        mainPanel(
            #input
            textInput("name", "What's your name?"),
           
             # Output: Formatted text for caption
            h3(textOutput("caption")),
        )
)

server <- function(input, output) {

    output$caption <-  renderText({
        paste0("Hello ", input$name)
    })
}

# Run the application 
shinyApp(ui = ui, server = server)
```
__2.8.2:__
```{r}
ui <- fluidPage(
    
    # Application title
    titlePanel("Chapter 2 Exercises"),
    
    # Main panel for displaying outputs 
    mainPanel(
        #input
        sliderInput("x", label = "If x is", min = 1, max = 50, value = 30),
        "then x times 5 is",
        
        h3(textOutput("product")),        
    )
)

server <- function(input, output, session) {
    output$product <- renderText({ 
        input$x * 5
    })
}

# Run the application 
shinyApp(ui = ui, server = server)
```
__2.8.3__
```{r}
ui <- fluidPage(
  
    # Application title
    titlePanel("Chapter 2 Exercises"),
  
    # Main panel for displaying outputs 
    mainPanel(
        #input
        sliderInput("x", label = "If x is", min = 1, max = 50, value = 30),
        sliderInput("y", label = "If y is", min = 1, max = 50, value = 5),
        strong("then x times y is"),
        
        h3(textOutput("product")),
        
    )
)

server <- function(input, output, session) {
    output$product <- renderText({ 
        input$x * input$y
    })
}

# Run the application 
shinyApp(ui = ui, server = server)
```
__2.8.4__
```{r}
library(shiny)

ui <- fluidPage(
  
  # Application title
    titlePanel("Chapter 2 Exercises"),
  
    sliderInput("x", "If x is", min = 1, max = 50, value = 30),
    sliderInput("y", "and y is", min = 1, max = 50, value = 5),
    "then, (x * y) is", textOutput("product"),
    "and, (x * y) + 5 is", textOutput("product_plus5"),
    "and (x * y) + 10 is", textOutput("product_plus10")
)

server <- function(input, output, session) {
    formulaText <- reactive(input$x * input$y)
        
    output$product <- renderText(formulaText())
    output$product_plus5 <- renderText(formulaText() + 5)
    output$product_plus10 <- renderText(formulaText() + 10)
}

# Run the application 
shinyApp(ui = ui, server = server)
```
__2.8.5__
This just required fixing three lines in this code block, which are annotated
```{r}
library(ggplot2)

datasets <- data(package = "ggplot2")$results[c(2, 4, 10), "Item"]

ui <- fluidPage(
  
  # Application title
    titlePanel("Chapter 2 Exercises"),
    
    selectInput("dataset", label = "Dataset", choices = datasets),
    verbatimTextOutput("summary"),
    
    #fixed: plotOutput instead of table output
    plotOutput("plot")
)

server <- function(input, output, session) {
    dataset <- reactive({
        get(input$dataset, "package:ggplot2")
    })
    
    #fixed: summary misspelled
    output$summary <- renderPrint({
        summary(dataset())
    })
    
    #fixed: need to call dataset as dataset()
    output$plot <- renderPlot({
        plot(dataset())
    }, res = 96)
}

shinyApp(ui = ui, server = server)
```

For [Chapter 3](https://mastering-shiny.org/basic-ui.html) I combined the collection of exercises into single shiny apps for each concept (inputs, outputs, layout). I've annotated the exercise number.

__Inputs__
```{r}
library(shiny)
ui <- fluidPage( 
    
    # Application title
    titlePanel("Chapter 3 Exercises"),
    
    # Main panel for displaying outputs 
    mainPanel(
        #3.2.8.1
        textInput("name", label=NULL,placeholder = "What's Your Name"),
        
        #3.2.8.2
        sliderInput("x", label = "When should we deliver"
                    , min = as.Date("2020-09-16"), max = as.Date("2020-09-23")
                    , value = as.Date("2020-09-17"), timeFormat = "%F", animate = FALSE),
        
        #3.2.8.3
        selectInput("animal", "What animals are you moving?", 
                    list(`Mammals` = list("Doggo", "Kitty", "IsThataHorse"),
                         `Reptiles` = list("Snek", "LizardWizard", "KingGizzard"),
                         `Rodent` = list("Gerbil", "Mouse", "Smooshy","MrPrickles"))
                    ,multiple = TRUE),
        
        #3.2.8.4
        sliderInput("y", label = "how much does it weigh"
                    , value = 150, min = 0, max = 1000, step = 5, animate = TRUE, interval),
        
        # Output: Formatted text for caption
        h3(textOutput("caption")),
        
    )
)


server <- function(input, output) {

    #this is just trying to make it semi cohesive/purposeful, not required, just for fun
    output$caption <-  renderText({
        paste0("Okay ", input$name
              ,", we will see you on ", input$x
              ,", to pick up your ", input$animal
              ,', weighing ', input$y, " pounds")
    })
}

# Run the application 
shinyApp(ui = ui, server = server)
```

__Outputs__
```{r}
ui <- fluidPage(
  
  # Application title
  titlePanel("Chapter 3 Exercises"),  
  
  plotOutput("plot", width = "400px"),
    dataTableOutput("table")
)
server <- function(input, output, session) {
    output$plot <- renderPlot(plot(1:5), 
                              #3.3.5.1
                              width = 700, height = 300, res = 96)
    
    output$table <- renderDataTable(mtcars, options = list(pageLength = 5, 
                              #3.3.5.2                             
                              ordering = FALSE, searching=FALSE))
}

# Run the application 
shinyApp(ui = ui, server = server)
```

__Layout__ 
```{r}
library(shinythemes)
ui <-  fluidPage(
    
    #3.4.6.3
    theme = shinytheme("sandstone"),
    
    titlePanel("Chapter 3 Exercises"),
    sidebarLayout(
        
        #3.4.6.2
        position = "right",
        sidebarPanel(
            numericInput("m", "Number of samples:", 2, min = 1, max = 100)
        ),
        
        #3.4.6.1
        mainPanel (
        fluidRow(
            column(10,
            plotOutput("hist")),
            
            column(10,
            plotOutput('scatter'))
        )
        )
    )
)


server <- function(input, output, session) {
    output$hist <- renderPlot({
        means <- replicate(1e4, mean(runif(input$m)))
        hist(means, breaks = 20)
    }, res = 96)
    
    #3.4.6.1
    output$scatter <- renderPlot(plot(1:5))
}
# Run the application 
shinyApp(ui = ui, server = server)
```

Feeling fairly confident, I read through [Chapter 4](https://mastering-shiny.org/basic-reactivity.html) and proceeded to the case study in [Chapter 5](https://mastering-shiny.org/basic-case-study.html#exercises-5).

```{r}
library(shiny)
library(vroom)
library(tidyverse)

#preprocesss
injuries <- vroom::vroom("Tamatoa/Exercises/Chapter5/neiss/injuries.tsv.gz")
products <- vroom::vroom("Tamatoa/Exercises/Chapter5/neiss/products.tsv")
population <- vroom::vroom("Tamatoa/Exercises/Chapter5/neiss/population.tsv")

count_top <- function(df, var, n) {
    df %>%
        mutate({{ var }} := fct_lump(fct_infreq({{ var }}), n = n)) %>%
        group_by({{ var }}) %>%
        summarise(n = as.integer(sum(weight)))
}

ui <- fluidPage(
    
    fluidRow(
        column(6,
               selectInput("code", "Product",
                           choices = setNames(products$prod_code, products$title),
                           width = "100%"
               )
        ),
        column(2, selectInput("y", "Y axis", c("rate", "count"))),
        
        column(4, sliderInput("x", label = "How many rows", min = 1, max = 25, value = 5))
    ),
    
    
    fluidRow(
        column(4, tableOutput("diag")),
        column(4, tableOutput("body_part")),
        column(4, tableOutput("location"))
    ),
    fluidRow(
        column(12, plotOutput("age_sex"))
    ),
    
    fluidRow(
        column(2, actionButton("story", "Tell me a story")),
        column(10, textOutput("narrative"))
    )
)

server <- function(input, output, session) {
    selected <- reactive(injuries %>% filter(prod_code == input$code))
    
    output$diag <- renderTable(count_top(selected(), diag, n = input$x), width = "100%")
    
    output$body_part <- renderTable(count_top(selected(), body_part, n = input$x), width = "100%")
    
    output$location <- renderTable(count_top(selected(), location, n = input$x), width = "100%")
    
    
    
    summary <- reactive({
        selected() %>%
            count(age, sex, wt = weight) %>%
            left_join(population, by = c("age", "sex")) %>%
            mutate(rate = n / population * 1e4)
    })
    
    output$age_sex <- renderPlot({
        if (input$y == "count") {
            summary() %>%
                ggplot(aes(age, n, colour = sex)) +
                geom_line() +
                labs(y = "Estimated number of injuries")
        } else {
            summary() %>%
                ggplot(aes(age, rate, colour = sex)) +
                geom_line(na.rm = TRUE) +
                labs(y = "Injuries per 10,000 people")
        }
    }, res = 96)
    
    output$narrative <- renderText({
        input$story
        selected() %>% pull(narrative) %>% sample(1)
    })
}

# Run the application 
shinyApp(ui = ui, server = server)
```

Needless to say, wow, there is a lot to make these things interesting and intelligent. I'm glad to play with Wickham's more established Shiny app, but also reminded that it's going to be a long road.

### Week Reflection, Coffee Project

Things are going well overall. 
I've been using Eight Ounce Coffee's (website)[https://eightouncecoffee.ca/] to start some scraping. My first week's scraper worked well, but wasn't rerunable in the second week. It seems there will be inconsitencies in the precise way they upload bags to the website -- mostly based on the number leading /n characters in the ```html_node``` character string. These /n characters are used as splitting mechanisms to grab the coffee name, rewrite it as a hypenated string, and then pushed into a for loop with ```read_html```. These individual hyperlinks to pull in the coffee details. When the data comes in, it's a bit messy, and it looks like some weeks will have more details available than others. I will have to build something flexible enough to properly write each metric to the appropriate column. Unlike last week, I wrote this week's coffee to a csv when it's all said and done so I don't have to rerun the HTML files again. Stuff gets taken down very quickly, as you'd expect. Once a week seems good though (having actually popped into The Library Specialty Coffee this week, their shelf mirrors Eight Ounce's website.) It shouldn't take long to grab a lot more roasters and use the same methodology. The resuling csv should be about 5 coffees per roaster, per week (maybe slower for some of the smaller roasters), so we'll have close to 100 coffees a week scraped off Eight Ounce. That's really nice.



